\documentclass[a4paper,11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{braket}

\numberwithin{equation}{section}
\renewcommand\thesubsection{\alph{subsection}}
\newcommand{\bvp}[1]{\mathbf{#1}'}
\newcommand{\bv}[1]{\mathbf{#1}}


%opening
\title{Statmech II HW2}
\author{Vince Baker}

\begin{document}

\maketitle

\section{Problem 1}
We prove the relation:
\begin{align}
 \ln{w(i)} = n_i\ln{(\frac{g_i}{n_i}-a)}-\frac{g_i}{a}\ln{(1-a\frac{n_i}{g_i})}
\end{align}
Where $a=-1$ for Bose-Einstein systems, $a=1$ for Fermi-Dirac systems, and $a=0$ for classical Maxwell-Boltzmann systems.
We assume that $n_i\gg 1$ and $g_i \gg 1$.\\
For Bose-Einstein systems:
\begin{align}
 w(i) &= \frac{(n_i+g_i-1)!}{(n_i)!(g_i-1)!}\\
 \ln{w(i)} &= \ln{\left(\frac{(n_i+g_i-1)!}{(n_i)!(g_i-1)!} \right)}\\
 \ln{w(i)} &= (n_i+g_i-1)\ln{(n_i+g_i-1)}-n_i\ln{n_i}-(g_i-1)\ln{(g_i-1)}\\
 \ln{w(i)} &= n_i\ln{\left(\frac{n_i+g_i-1}{n_i}\right)} +(g_i-1)\ln{(\frac{n_i}{g_i+1}+1)}\\
 \ln{w(i)} &= n_i\ln{\left(\frac{g_i}{n_i}+1\right)} +g_i\ln{(\frac{n_i}{g_i}+1)}
\end{align}
Where in 1.5 we have used $n_i\gg 1$ and $g_i \gg 1$ to simplify the expression.
\\
For Fermi-Dirac systems:
\begin{align}
  w(i) &= \frac{(g_i)!}{(n_i)!(g_i-n_i)!}\\
 \ln{w(i)} &= \ln{\left(\frac{(g_i)!}{(n_i)!(g_i-n_i)!} \right)}\\
 \ln{w(i)} &= g_i\ln{g_i}-n_i\ln{n_i}-(g_i-n_i)\ln{(n_i-g_i)}\\
 \ln{w(i)} &= n_i\left(\ln{(g_i-n_i)}-\ln{n_i} \right)+g_i\left(\ln{\frac{g_i}{g_i-n_i}} \right)\\
 \ln{w(i)} &= n_i\ln{\left(\frac{g_i}{n_i}-1 \right)}-g_i\ln{\left(1-\frac{n_i}{g_i} \right)}
\end{align}
\\
For Maxwell-Boltazmann systems:
\begin{align}
  w(i) &= \frac{(g_i)^{n_i}}{(n_i)!}\\
 \ln{w(i)} &= n_i\ln{g_i}-n_i\ln{n_i}+n_i
\end{align}
Taking the limit of $-\frac{g_i}{a}\ln{(1-a\frac{n_i}{g_i})} $ as $a \rightarrow 0$ using L'Hopital's rule:
\begin{align}
 \varinjlim 0 -\frac{g_i}{a}\ln{(1-a\frac{n_i}{g_i})} &= n_i\\
 \ln{w(i)} &= n_i\ln{g_i}-n_i\ln{n_i}-\frac{g_i}{a}\ln{(1-a\frac{n_i}{g_i})}\ (a=0)
\end{align}
\\
\section{Problem 2}
We prove the maximization result using Lagrange multipliers:
\begin{align}
 \delta \ln{W_i}-\left(\alpha\sum \delta n_i+\beta\sum \epsilon_i \delta n_i \right) &= \sum \left(\ln{(\frac{g_i}{n_i}-a)}-\alpha-\beta \epsilon_i \right)\delta n_i
\end{align}
We take the derivative with respect to $n_i$ of $\ln{w_i}$ (equation 1.1).
\begin{align}
 \frac{\partial \ln{w_i}}{\partial n_i} &= \ln{(\frac{g_i}{n_i}-a)}-\frac{g_i}{n_i}\frac{1}{g_i/n_i-a}+\frac{1}{1-an_i/g_i}\\
 \frac{\partial \ln{w_i}}{\partial n_i} &= \ln{(\frac{g_i}{n_i}-a)}
\end{align}
With $W_i=\prod w_i$ and using the logarithm addition rules this proves 2.1.
\
\section{Problem 3}
Starting from the grand canonical equivalent of the microcanonical ensmble expression for number of microstates, with $g_i=1$ and replacing $n_i^*$ with $<n_e>$:
\begin{align}
 \ln{W} &= \sum_e\left(<n_e>\ln{(\frac{1}{<n_e>}-a)}-\frac{1}{a}\ln{(1-a<n_e>)} \right)\\
 S & =k\ln{W} = k \sum_e\left(<n_e>\ln{(\frac{1}{<n_e>}-a)}-\frac{1}{a}\ln{(1-a<n_e>)} \right)\\
 S &= k \sum_e <n_e>\ln{\left(\frac{1-a<n_e>}{<n_e>} \right)}-\frac{1}{a}\ln{(1-a<n_e>)}\\
 S &= k\sum_e \left(<n_e>-\frac{1}{a}\right)\ln{(1-a<n_e>)}-<n_e>\ln{<n_e>}
\end{align}
So we have proved the relation with $a=-1$ for Bosons and $a=1$ for Fermions.
\\
For Fermions there are only two probabilities:
\begin{align}
 \rho_0 = 1-<n_e>\\
 \rho_1 = <n_e>
\end{align}
So the probability form of the entropy is:
\begin{align}
 S &= -k\sum_e \left((1-<n_e>\ln{(1-<n_e>)}+<n_e>\ln{<n_e>}  \right)\\
 S &= k\sum_e \left((<n_e>-1)\ln{(1-<n_e>)}-<n_e>\ln{<n_e>}  \right)
\end{align}
\\
For Bosons, the probability is:
\begin{align}
 \rho_e = \frac{<n_e>^n}{\left(<n_e>+1 \right)^{n+1}}
\end{align}
We can write the entropy equation terms as the expectation value of $\ln{\rho_e}$.
\begin{align}
 \sum_n \rho_e\ln{\rho_e} &= <\ln{\rho_e}>\\
 S &= -k\sum_e \braket{\ln{\braket{\rho_e}}}\\
 S &= -k \sum_e \braket{n\ln{\braket{n_e}}-(n+1)\ln{(\braket{n_e}+1)}}\\ 
 S &= k \sum_e (\braket{n_e}+1)\ln{(\braket{n_e}+1)}-\braket{n_e}\ln{\braket{n_e}}
\end{align}

\end{document}
